\pdfbookmark[0]{ABSTRACT}{abstract}
\chapter*{ABSTRACT}

\indent 

In this thesis, we apply a recently developed technique to
comprehensively categorize all possible families of solitary wave
solutions in two models of topical interest.

The models considered are:
\begin{itemize}
\item the Generalized Pochhammer-Chree Equations, which  govern the propagation of longitudinal waves in elastic rods,

and

\item a generalized microstructure PDE.
\end{itemize}

Limited analytic results exist for the occurrence of one family of  solitary
wave solutions for each of these equations. Since, as mentioned above, solitary
wave solutions often play a central role in the long-time evolution of an
initial disturbance, we consider such solutions of both models here (via the
normal form approach) within the framework of reversible systems theory.

Besides confirming the existence of the known family of solitary waves for each
model, we find a continuum of delocalized solitary waves (or homoclinics to
small-amplitude periodic orbits).  On isolated curves in the relevant parameter
region, the delocalized waves reduce to genuine embedded solitons.  For the
Microstructure equation, the new family of solutions occur in regions of
parameter space distinct from the known solitary wave solutions and are thus
entirely new.

Directions for future work, including the dynamics of each family of
solitary waves using exponential asymptotics techniques, are also mentioned.

%%%%%  DEDICATION SECTION -- OPTIONAL %%%%%

%\newpage
%\rule{0pt}{0pt}
%\begin{center}
%\vspace{\stretch{.8}}

\dedication{
Dedicated To My Family
}

%\vspace{\stretch{1.2}}
%\end{center}

%%%%%  END OF DEDICATION SECTION  %%%%%

\newpage
\pdfbookmark[0]{ACKNOWLEDGMENTS}{acknowledgments}
\chapter*{ACKNOWLEDGMENTS}

I would like to thank Dr. Roy Choudhury, Dr. Ram Mohapatra and Dr. Constance Schobert for their support and guidance as committee members.

\newpage
